Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow

### 2009-05-20 paper/paper.tex
Polishing of paper - cleaned up code to remove warnings.

Managing evidence in a globally distributed system requires the use of
globally unique identifiers to ensure no name collisions can occur
with disconnected systems. AFF1 assigns each piece of evidence a
unique 128-bit identifier called a GID but did not make it clear when
this identifier should be changed and when it should remain the same.

@@

In this case they can each create a new volume which extends to
original volume and save their analysis on this new volume. Now they
only need to share this new volume with other analysts who also have a
copy of the old volume to interchange their findings.

This is made possible because each volume is independent of one
another, but can still be recombined into a bigger evidence set.

@@

\item \emph{A Target Reference} is a way of referencing objects by use
  of a Uniform Resource Identifier (URI). The URI can be another AFF
  Object URN or may be a more general Uniform Resource Locator (URL),
  such as for example a HTTP or FTP object. This innovation allows
  objects in one volume to refer to objects in different volumes,
  facilitating data fusion and cross referencing.

@@

The AFF4 design adopts a scheme of globally unique identifiers for
identifying and referring to all evidence. We define an AFF4 specific
URN scheme, which we call the AFF4 URN. URN's of this scheme use the
namespace\cite{RFC1737} ``\emph{aff4}'' and therefore begin with the 
string(\emph{urn:aff4}). AFF4 URNs are then be made unique by use of 
a unique identifier generated
as per RFC4122\cite{RFC4122}. For example, an AFF4 URN might be
\emph{urn:aff4:bcc02ea5-eeb3-40ce-90cf-7315daf2505e}

@@

It is important to stress that the properties file is simply a
serialization of statements into volume segments. The statements may
exist without being stored in a volume (for example, being stored on
an external SQL server). Alternatively, these statements may be stored
in some other way inside or outside the volume (e.g. SQLite
databases).

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visibile to it. AFF4 provides a mechanism to use an
external resolver as well---for example, we have implemented a
resolver that stores Attributes in a MySQL database to provide for a
Universal Resolver that shares information between different instances
on the same network.

Each URN within the AFF4 universe must have an ``\texttt{aff4:type}''
attribute to denote the type of the Object.

@@

\section{Volumes}
The volume object is responsible for providing storage for
segments. Segments are stored and retrieved using their URNs. We
describe two different implementations of volume objects, namely the
{\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
convert from one implementation to another easily, without affecting
any external references. This is because external references are to
the Volume URN itself, so it is safe to convert from one
representation to another. In this way its possible for example, to
unzip a Zip64 volume into a directory and vice versa.

It is important to emphasise that Volumes are merely containers which
provide storage for segments. There is no restriction of which
segments can be stored by any particular volume. For example, the
segments which make up a single Image stream may be stored in a number
of volumes (splitting the image in some way among them). Similarly,
the segments representing a number of streams may be stored in the
same volume.

@@

The Directory Volume uses the {\em aff4:stored} attribute to provide a
base URL. The URL for each segment is then constructed by appending
the escaped segment URN to the base URL. Note that there is no
restriction on what type of URL this can be, so it may be a location
on a filesystem (e.g. {\em file://some\_directory/}) or a location on a
HTTP server (e.g. {\em http://intranet.server/some\_path}). In this
way its possible to move the entire volume from a filesystem to a web
server transparently.

@@

\item Zip64 already includes support for the ``table of contents''
  envisioned but never implemented by Garfinkel; Zip64 calls this
  table the ``Central Directory.'' It is written at the end of the
  Zip64 volume.

@@

Figure~\ref{zip_structure} shows the basic structure of a Zip
archive. As can be seen, the archive consists of a {\em Central Directory} (CD)
located at the end of the archive. The CD is a list of pointers to
individual {\em File header} structures located within the body of the
archive. Headers are then followed by the file data, after it has been
compressed by the appropriate compression method (as specified in the
header). Each archived file is optionally followed by a {\em Data
Descriptor} describing the length and CRC of the archived file. Using
the data descriptor field allows implementations to write archives
without needing to seek in the output file. This allows Zip files to
be written to pipes for example, sending an image over the network
using netcat or ssh.

@@

It is important to note that AFF4 only requires that the volume be
capable of storing multiple named segments of data. Although our AFF4
implementation uses the Zip64 file format as an underlying storage
mechanism, our system also supports legacy AFF1 volumes as well as
segments that are stored in other storage systems---for example,
Amazon.com's Simple Storage System (S3)\cite{s3-aws-home-page-money}.

@@

\item Simple Zip implementations might rescan the
  Central Directory for each segment request. Since in practice there
  can be a large number of segments in a volume, it is advisable to
  have a Zip64 implementation that is optimized to storing thousands
  (or even hundreds of thousands) of segments in an efficient data
  structure. In fact our implementation uses the Universal Resolver to
  store the parsed central directory information, which means that in
  most cases we dont even need to scan the Central Directory at all.

\item While the Zip specification duplicates data found in the Central
  Directory entry in each File Header (such as filename, size, CRC
  etc), many implementations that we have examined only populate this
  information in one of these places. In the interest of robustness,
  we wanted to ensure that data stored in both locations would be
  populated to allow recovery of at least \emph{some} evidence that
  might exist in damaged volumes. If the central directory is lost, it
  is possible to scan through the volume, and locate all the Zip64
  file headers. Then its possible to repair and reconstruct the
  central directory.

@@

Image streams specify the \texttt{chunk\_size} attribute, as the
number of image bytes each chunk contains (typically chunk sizes are
32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
which specifies how many chunks are stored in each bevy. Each chunk is
compressed individually using the zlib compress algorithm. This
general structure of storing chunks within larger segments is similar
to the technique used by the Expert Witness file format (EWF) used by
EnCase\cite{encase-3.0} and implemented by the open source
libewf\cite{libewf} package. This improvement from AFF1's 16MB segment
size results in a better match between requested size and the minimum
size required for decompression. This means that less data is needed
to be decompressed unnecessarily where reading small sectors randomly.

@@

\subsection{The Map Stream}
\label{map_stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image, while a TCP/IP stream is simply a collection of
payloads from selected network packets. Sometimes the same data may be
viewed in a number of ways---for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation.  Zero Storage Carving\cite{Meijer2006} is a way of
specifying carved files in terms of a sequence of blocks taken from
the image; Cohen extended this concept to an arbitrary mapping
function\cite{1363239,Cohen2007} which can be used to describe
arbitrary mappings of carved files within a single image.

@@

Denoting the stream offset by $x$, and the target offset by $y$, the
Map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
byte at a mapped stream offset $x$ can then be satisfied by reading a
byte from target $T_i$ at offset $y$ given by:
\begin{eqnarray}
y = (x - X_i) + Y_i & &
\forall x \in \left [X_i, X_i+1 \right )
\end{eqnarray}

@@

In order to efficiently express periodic maps such as those found in
RAID arrays, the Map stream may be provided with two optional
parameters a {\em target\_period} ($T_p$), and {\em stream\_period}
($S_p$). If specified, the above relation becomes:
\begin{eqnarray*}
p &:=& floor\left (\frac{x}{S_p} \right) \\
x' &:=& mod(x ,S_p)  \\   \label{eq:no1}
y &:=& (x'-X_i) + Y_i + p \times T_p
\end{eqnarray*}

@@

For this reason, URLs may be used interchangeably with a URN within
the AFF4 universe. For example, the \emph{aff4:stored} attribute of a
volume may be specified as a URL
(e.g. \emph{http://intranet/123453/}).  AFF4 provides transparent
support for HTTP and FTP URLs by means of the curl HTTP
library\cite{libcurl}. The HTTP Stream, therefore satisfies read
requests by making HTTP requests to the web server. We use the
\emph{Content-Range} HTTP header to request exactly the byte range the
client is interested in. This allows efficient network transport as we
do not need to download unnecessary data, we just request those chunks
the client application requires.

Our implementation also enables direct writing to a HTTP URL using the
WebDav extensions to HTTP\cite{webdav-rfc}. The HTTP stream also
supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
Layer---SSL) protocols transparently, as provided by the curl library.

@@

The result is that a number of AFF4 volumes are used as {\em Container
Volumes} to provide storage for Encrypted Streams. The main {\em
Embedded Volume}, which actually contains data is stored within the
Encrypted Stream, effectively distributed throughout the container
volumes. Note that the outer Volume may contain several Encrypted
Streams and therefore contain multiple AFF4 Encrypted
Volumes. Container Volumes may contain non encrypted streams as well,
and may implement different encryption schemes and keys for each
Encrypted stream. This effectively allows arbitrary access policies to
be implemented as only volumes which can be accessed can be merged.

@@

An AFF4 Identity object represents an entity, currently described by
way of an X509 certificate. The URN of an identity object is the
certificate's fingerprint, and is therefore unique to the
certificate. Identity objects contains \emph{aff4:statement}
attributes which refer to AFF4 streams containing statements. The
identity object also contains a copy of the certificate used to sign
the statements.

@@

\subsection{Using distributed evidence}
An investigation is conducted by the New York team. The case relates
to a hard disk image stream stored inside a volume, in turn stored on
the NY evidence server at URL http://ny.wan/evidence1.aff4 . The team
requires an analyst (Bob) in LA to assist with their analysis. The LA
analyst types\footnote{fls is the file listing command which is part
of the Sleuthkit}:

@@

It is still advantageous in this case to bring discrete files into the
AFF4 evidence universe by acquiring each evidence file using a unique
URN. As explained in \ref{definitions}, segments are AFF4 stream
objects which are implemented by storing the data in Zip archive
members. It follows, therefore, that a regular zip file containing
files is also a valid AFF4 volume.

@@

Alice uses windows explorer to obtain a Zip file of the files of
interest. After taking the file back to the lab, she then signs the
files, and adds a volume URN, making the Zip file a fully compliant
AFF4 volume.

>>>>

Managing evidence in a globally distributed system requires the use of
globally unique identifiers to ensure no name collisions can occur
with disconnected locations. AFF1 assigns each piece of evidence a
unique 128-bit identifier called a GID but did not make it clear when
this identifier should be changed and when it should remain the same.

@@

In this case they can each create a new volume which extends the
original volume and save their analysis on this new volume. Now they
only need to share this new volume with other analysts who also have a
copy of the old volume to interchange their findings.

This is made possible because each volume is independent of one
another, but is still viewed as part of a bigger evidence set.

@@

\item \emph{A Reference} is a way of referencing objects by use
  of a Uniform Resource Identifier (URI). The URI can be another AFF
  Object URN or may be a more general Uniform Resource Locator (URL),
  such as for example a HTTP or FTP object. This innovation allows
  objects in one volume to refer to objects in different volumes,
  facilitating data fusion and cross referencing.

@@

The AFF4 design adopts a scheme of globally unique identifiers for
identifying and referring to all evidence. We define an AFF4 specific
URN scheme, which we call the AFF4 URN. URN's of this scheme use the
namespace\cite{RFC1737} ``\emph{aff4}'' and therefore begin with the 
string ``\emph{urn:aff4}''. AFF4 URNs are then be made unique by use of 
a unique identifier generated
as per RFC4122\cite{RFC4122}. For example, an AFF4 URN might be
\emph{urn:aff4:bcc02ea5-eeb3-40ce-90cf-7315daf2505e}

@@

It is important to stress that the properties file is simply a
serialization of statements into volume segments. The statements may
exist without being stored in a volume (for example, being stored on
an external SQL server). Alternatively, these statements may be stored
in some other way inside or outside the volume (e.g. SQLite
database files).

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visible to it. AFF4 provides a mechanism to use an
external resolver as well---for example, we have implemented a
resolver that stores Attributes in a MySQL database to provide for a
persistent Universal Resolver that shares information between
different instances on the same network.

Each URN within the AFF4 universe must have an ``\texttt{aff4:type}''
attribute to denote the type of the Object. Objects may also have a
the ``\texttt{aff4:interface}'' attribute to denote what kind of
interface they present (e.g. {\emph stream} or {\emph volume}).

@@

\section{Volumes}
The volume object is responsible for providing storage for
segments. Segments are stored and retrieved using their URNs. We
describe two different implementations of volume objects, namely the
{\em Directory Volume} and the {\em ZipFile Volume}. It is possible to
convert from one implementation to another easily, without affecting
any external references.

It is important to emphasize that Volumes are merely containers which
provide storage for segments. There is no restriction of which
segments can be stored by any particular volume. For example, the
segments which make up a single Image stream may be stored in a number
of volumes (splitting the image in some way among them). Similarly,
the segments representing a number of streams may be stored in the
same volume.

@@

The Directory Volume uses the {\em aff4:stored} attribute to provide a
base URL. The URL for each segment is then constructed by appending
the escaped segment URN to the base URL. Note that there is no
restriction on what type of URL this can be, so it may be a location
on a filesystem (e.g. {\em file:///some/directory/}) or a location on a
HTTP server (e.g. {\em http://intranet.server/some/path}). In this
way its possible to move the entire volume from a filesystem to a web
server transparently.

@@

@@

Figure~\ref{zip_structure} shows the basic structure of a Zip
archive. As can be seen, the archive consists of a {\em Central Directory} (CD)
located at the end of the archive. The CD is a list of pointers to
individual {\em File header} structures located within the body of the
archive. Headers are then followed by the file data, after it has been
compressed by the appropriate compression method (as specified in the
header). Each archived file is optionally followed by a {\em Data
Descriptor} describing the length and CRC of the archived file. Using
the data descriptor field allows implementations to write archives
without needing to seek in the output file. This allows Zip files to
be written to pipes for example, sending an image over the network
using netcat or ssh. AFF4 always uses the data description header to
ensure volumes are written continuously without needing to seek in the
output file.

@@

It is important to note that AFF4 only requires that the volume be
capable of storing multiple named segments of data. Although our AFF4
implementation uses the Zip64 file format as an underlying storage
mechanism, our system also supports legacy AFF1 volumes as well as
Expert Witness Evidence files \cite{libewf}.

@@

\item Simple Zip implementations might rescan the
  Central Directory for each segment request. Since in practice there
  can be a large number of segments in a volume, it is advisable to
  have a Zip64 implementation that is optimized to storing thousands
  (or even hundreds of thousands) of segments in an efficient data
  structure. In fact our implementation uses the Universal Resolver
  itself to store the parsed central directory information, which
  means that in most cases we do not even need to scan the Central
  Directory at all.

\item While the Zip specification duplicates data found in the Central
  Directory entry in each File Header (such as filename, size, CRC
  etc), many implementations that we have examined only populate this
  information in one of these places. In the interest of robustness,
  we wanted to ensure that data stored in both locations would be
  populated to allow recovery of at least \emph{some} evidence that
  might exist in damaged volumes. If the central directory is lost, it
  is possible to scan through the volume, and locate all the Zip64
  file headers. Then it is possible to repair and reconstruct the
  central directory.

@@

Image streams specify the \texttt{chunk\_size} attribute, as the
number of image bytes each chunk contains (chunk size defaults to
32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
which specifies how many chunks are stored in each bevy. Each chunk is
compressed individually using the zlib compress algorithm. This
general structure of storing chunks within larger segments is similar
to the technique used by the Expert Witness file format (EWF) used by
EnCase\cite{encase-3.0} and implemented by the open source
libewf\cite{libewf} package. This improvement from AFF1's 16MB segment
size results in a better match between requested size and the minimum
size required for decompression. Less data is needed to be
decompressed unnecessarily where reading small sectors randomly,
leading to vast performance improvements.

@@

\subsection{The Map Stream}
\label{map_stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image, while a TCP/IP stream is simply a collection of
payloads from selected network packets. Sometimes the same data may be
viewed in a number of ways---for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation \cite{Tanenbaum2008}.  Zero Storage
Carving\cite{Meijer2006} is a way of specifying carved files in terms
of a sequence of blocks taken from the image; Cohen extended this
concept to an arbitrary mapping function\cite{1363239,Cohen2007} which
can be used to describe arbitrary mappings of carved files within a
single image.

@@

Denoting the stream offset by $x$, and the target offset by $y$, the
Map specifies a set of points $(X_i,Y_i,T_i)$. Read requests for a
byte at a mapped stream offset $x$ can then be satisfied by reading a
byte from target $T_i$ at offset $y$ given by:
\begin{eqnarray}
y = (x - X_i) + Y_i & &
\forall x \in \left [X_i, X_{i+1} \right )
\end{eqnarray}

@@

In order to efficiently express periodic maps such as those found in
RAID arrays, the Map stream may be provided with two optional
parameters: a {\em target\_period} ($T_p$), and {\em stream\_period}
($S_p$). If specified, the above relation becomes:
\begin{eqnarray*}
p &:=& floor\left (\frac{x}{S_p} \right) \\
x' &:=& mod(x ,S_p)  \\   \label{eq:no1}
y &:=& (x'-X_i) + Y_i + p \times T_p
\end{eqnarray*}

@@

For this reason, URLs may be used interchangeably with a URN within
the AFF4 universe. For example, the \emph{aff4:stored} attribute of a
volume may be specified as a URL
(e.g. \emph{http://intranet/123453/}).  AFF4 provides transparent
support for HTTP and FTP URLs by means of the Curl HTTP
library\cite{libcurl}. The HTTP Stream, therefore satisfies read
requests by making HTTP requests to the web server. We use the
\emph{Content-Range} HTTP header to request exactly the byte range the
client is interested in. This allows efficient network transport as we
do not need to download unnecessary data, we just request those chunks
the client application requires.

Our implementation also enables direct writing to a HTTP URL using the
WebDav extensions to HTTP\cite{webdav-rfc}. The HTTP stream also
supports the File Transfer Protocol (FTP) and HTTPS (Secure Sockets
Layer---SSL) protocols transparently, as provided by the Curl library.

@@

The result is that a number of AFF4 volumes are used as {\em Container
Volumes} to provide storage for Encrypted Streams. The main {\em
Embedded Volume}, which actually contains data is stored within the
Encrypted Stream, effectively distributed throughout the container
volumes. Note that the outer Volume may contain several Encrypted
Streams and therefore contain multiple AFF4 Encrypted
Volumes. Container Volumes may contain non encrypted streams as well,
and may implement different encryption schemes and keys for each
Encrypted stream. This effectively allows arbitrary access policies to
be implemented as only volumes which can be accessed can be read.

@@

An AFF4 Identity object represents an entity, currently described by
way of an X509 certificate. The URN of an identity object is the
certificate's fingerprint, and is therefore unique to the
certificate. Identity objects contain \emph{aff4:statement}
attributes which refer to AFF4 streams containing statements. The
identity object also contains a copy of the certificate used to sign
the statements.

@@

\subsection{Using distributed evidence}
An investigation is conducted by the New York team. The case relates
to a hard disk {\emph Image stream} stored inside a volume, in turn
stored on the NY evidence server at URL http://ny.wan/evidence1.aff4
. The team requires an analyst (Bob) in LA to assist with their
analysis. The LA analyst types\footnote{fls is the file listing
command which is part of the Sleuthkit}:

@@

It is still advantageous in this case to bring discrete files into the
AFF4 evidence universe by acquiring each evidence file using a unique
URN. As explained in Section \ref{definitions}, segments are AFF4 stream
objects which are implemented by storing the data in Zip archive
members. It follows, therefore, that a regular zip file containing
files is also a valid AFF4 volume.

@@

Alice uses windows explorer to obtain a Zip file of the files of
interest. After taking the archive back to the lab, she then signs the
files, and adds a volume URN, making the Zip file a fully compliant
AFF4 volume.
